### 2024-03-27 course/csic5011/2024/project/project.tex
lecture 07

\graphicspath{{figures/}{../book.datasci/figures/}{../mySlides/figures/}}

@@

\lecture{Final Project}{Yuan Yao}{Due: 23:59 18 May, 2023}{24 Apr, 2023}

@@

\begin{enumerate}
\item Pick up ONE (or more if you like) favourite dataset below to work. If you would like to work on a different problem outside the candidates we proposed, please email course instructor about your proposal.  
\item Team work: we encourage you to form small team, up to FOUR persons per group, to work on the same problem. Each team must submit:
\subitem(a) \emph{ONE report}, \emph{with a clear remark on each person's contribution}. The report can be in the format of a \emph{technical report within 8 pages}, e.g. NIPS conference style 
\begin{center}
\url{https://nips.cc/Conferences/2016/PaperInformation/StyleFiles} \\
\noindent and a sample file at \\
\url{https://arxiv.org/pdf/1606.04930.pdf} 
\end{center}
or of a \emph{poster}, e.g. 
\begin{center}%\url{http://math.stanford.edu/~yuany/publications/poster_CleaveBioCPH2017_ForReview.pptx}
\url{https://github.com/yuany-pku/2017_math6380/blob/master/project1/DongLoXia_poster.pptx}
\end{center}
\subitem(b) \emph{ONE short presentation video within 10 mins}, e.g. in Youtube or Bilibili link. You may submit your presentation slides together with the video link to help understanding. 
\item In the report, (1) design or raise your scientific problems (a good problem is sometimes more important than solving it); (2) show your main results with a careful analysis supporting the results toward answering your problems. Remember: scientific analysis and reasoning are more important than merely the performance results. Reproducible source codes may be submitted via a \url{https://github.com/} link, through email as a zip file, or as an appendix of the report if it is not large. 
\item Upon the start, submit your grouping information (representer/member) to the following address (\href{mailto:datascience.hw@gmail.com}{datascience.hw@gmail.com}) with Title: \underline{CSIC 5011: Project 2 [grouping]}. Then we shall create public folders and you should submit your report to your group folder under subfolder \texttt{[GROUP]/report}. % (\href{mailto:datascience\_hw@126.com}{datascience\_hw@126.com}). 
\end{enumerate}

\newpage

@@

\section{PageRank and Primary Eigenvectors}

\section{Transition Paths of Karate Club Network} 
The following dataset contains a 34-by-34 adjacency matrix $A$ of Zachery's Karate Club Network.

\url{https://github.com/yao-lab/yao-lab.github.io/blob/master/data/karate.mat}

As shown in Figure \ref{fig:karate}, node 1 represents the coach of the club and node 34 is the owner (president) of the club. The undirected, unweighted edges between nodes represent
the affinity relation between club members. The story behind the network is this: the coach would like to raise the instruction fee while the president does not allow this; the conflicts finally result in a fission of the club -- the coach leaves the club with his funs and sets up his own club marked in red, and the blue nodes remain in the old club with the president.  
\begin{figure}[h]
\centering
\includegraphics[width=0.3\textwidth]{karate1-a_old.eps}
\caption{Zachery's Karate Club Network}\label{fig:karate}
\end{figure}

A. Apply the spectral clustering via the Cheeger vector (the second smallest eigenvector associated with normalized Graph Laplacian) to bipartite the network into two components, and compare it with the ground truth fission above.

B. Perform the following experiment with the transition path analysis. 
\begin{enumerate}
\item Define a Markov chain according to the network structure, such that from each node a random walker will jump to its neighbors with equal probability, i.e. $P=D^{-1} A$ where $D=\diag(d_i)$ and $d_i = \sum_j A_{ij}$;
\item Compute its stationary distribution $\pi(i)\sim d_i$. 
\item Define the source set $V_0=\{1\}$ and the target set $V_1=\{34\}$, compute the committor function
\[  q(x) = Prob(\mbox{trajectory starting from $x$ hitting $V_1$ before $V_0$}) \]
by solving the following Dirichlet boundary problem
\[ (L q)(x) = 0, \ \ \ x\in V_u := V -\{1,34\}, \ \ \ q(1)=0,\ q(34)=1. \]
Find those edges which contains one node $q(x)\leq 0.5$ and the other node $q(x)\geq 0.5$. Such edge set defines a cut of the graph. 
\item Compute the effective flux on each edge $(x,y)$ by 
\[ J^+(x,y) = \max(J(x,y) - J(y,x), 0 ),\]
where
\[     J(x,y) = 
     \begin{cases}
       \pi(x) (1-q(x)) P_{xy}q(y), & x\neq y;\\
       0, & \text{otherwise}.
     \end{cases}
\]
\item Compute the transition flux through each node $x\in V$ by 
\[
     T(x) = \left\{\begin{array}{ll}
     \sum_{y \in V} J^+(x,y) ,& x\in V_0 \\
     \sum_{y\in V} J^+(y,x), & x\in V_1 \\
     \sum_{y\in V} J^+(x,y) = \sum_{y\in V} J^+(y,x), & x\in V_u 
     \end{array}\right.
\]
\item Visualize your results by plotting a directed graph, with an arrow on each edge indicating the effective flux direction $J^+(x,y)>0$, different color marking the cut set of the graph, and if possible edge/node size in proportion to the size of effective/transition flux.
\end{enumerate}

A reference can be seen at:
\begin{itemize}
\item Weinan E, Jianfeng Lu, and Yuan Yao. {\it The Landscape of Complex Networks: Critical Nodes and A Hierarchical Decomposition}. Methods and Applications of Analysis, special issue in honor of Professor Stanley Osher on his 70th birthday, 20(4):383-404, 2013. \\ arXiv: \url{http://arxiv.org/abs/1204.6376}. \\
Pdf Link: \url{https://github.com/yao-lab/yao-lab.github.io/blob/master/reference/ELY.MAA13.pdf} 
\end{itemize}

The following matlab codes implement the transition path analysis and reproduce the results in the paper above:

\url{https://github.com/yao-lab/yao-lab.github.io/blob/master/data/karate_tpt.m}

You may explore more networks, including the two other examples in the paper above:

LAO-binding network (provided by Xuhui Huang): \url{https://github.com/yao-lab/yao-lab.github.io/blob/master/data/lao54.mat}

Les Miserables social network: \url{https://github.com/yao-lab/yao-lab.github.io/blob/master/data/lesmis.mat}, or \url{https://github.com/yao-lab/yao-lab.github.io/blob/master/data/lesmis.txt}

@@

\section{Hand-written Digits} The website

@@

\section{Finance Data}
The following data contains 1258-by-452 matrix with closed prices of 452 stocks in SNP'500 for workdays in 4 years.

@@

\section{Animal Sleeping Data} The following data contains animal sleeping hours together with other features:

@@

\section{US Crime Data} The following data contains crime rates in 59 US cities during 1970-1992:

@@

>>>>

\graphicspath{{figures/}}

@@

\lecture{Final Project}{Yuan Yao}{Due: 23:59 30 April, 2024}{27 Mar, 2024}

@@

\begin{enumerate}
\item Pick up ONE (or more if you like) favourite dataset below to work. If you would like to work on a different problem outside the candidates we proposed, please email course instructor about your proposal.  
\item Team work: we encourage you to form small team, up to FOUR persons per group, to work on the same problem. Each team must submit:
\subitem(a) \emph{ONE report}, \emph{with a clear remark on each person's contribution}. The report can be in the format of a \emph{technical report within 8 pages}, e.g. NIPS conference style 
\begin{center}
\url{https://nips.cc/Conferences/2016/PaperInformation/StyleFiles} \\
\noindent and a sample file at \\
\url{https://arxiv.org/pdf/1606.04930.pdf} 
\end{center}
or of a \emph{poster}, e.g. 
\begin{center}%\url{http://math.stanford.edu/~yuany/publications/poster_CleaveBioCPH2017_ForReview.pptx}
\url{https://github.com/yuany-pku/2017_math6380/blob/master/project1/DongLoXia_poster.pptx}
\end{center}
\subitem*(b)\footnote{We plan to do final project presentation in the last lecture.} \emph{ONE short presentation video within 10 mins}, e.g. in Youtube or Bilibili link. You may submit your presentation slides together with the vieo link to help understanding. 
\item In the report, (1) design or raise your scientific problems (a good problem is sometimes more important than solving it); (2) show your main results with a careful analysis supporting the results toward answering your problems. Remember: scientific analysis and reasoning are more important than merely the performance results. Reproducible source codes may be submitted via a \url{https://github.com/} link, through email as a zip file, or as an appendix of the report if it is not large. 
\item Upon the start, submit your grouping information (representer/member) to the following address (\href{mailto:datascience.hw@gmail.com}{datascience.hw@gmail.com}) with Title: \underline{CSIC 5011: Project 2 [grouping]}. Then we shall create public folders and you should submit your report to your group folder under subfolder \texttt{[GROUP]/report}. % (\href{mailto:datascience\_hw@126.com}{datascience\_hw@126.com}). 
\end{enumerate}

@@

@@

\subsection{Hand-written Digits} The website

@@

\subsection{Finance Data}
The following data contains 1258-by-452 matrix with closed prices of 452 stocks in SNP'500 for workdays in 4 years.

@@

\subsection{Animal Sleeping Data} The following data contains animal sleeping hours together with other features:

@@

\subsection{US Crime Data} The following data contains crime rates in 59 US cities during 1970-1992:

@@

%
%\section{Transition Paths of Karate Club Network} 
%The following dataset contains a 34-by-34 adjacency matrix $A$ of Zachery's Karate Club Network. 
%
%\url{https://github.com/yao-lab/yao-lab.github.io/blob/master/data/karate.mat}
%
%As shown in Figure \ref{fig:karate}, node 1 represents the coach of the club and node 34 is the owner (president) of the club. The undirected, unweighted edges between nodes represent
%the affinity relation between club members. The story behind the network is this: the coach would like to raise the instruction fee while the president does not allow this; the conflicts finally result in a fission of the club -- the coach leaves the club with his funs and sets up his own club marked in red, and the blue nodes remain in the old club with the president.  
%\begin{figure}[h]
%\centering
%\includegraphics[width=0.3\textwidth]{karate-a.pdf}
%\caption{Zachery's Karate Club Network}\label{fig:karate}
%\end{figure}
%
%A. Apply the spectral clustering via the Cheeger vector (the second smallest eigenvector associated with normalized Graph Laplacian) to bipartite the network into two components, and compare it with the ground truth fission above.  
%
%B. Perform the following experiment with the transition path analysis. 
%\begin{enumerate}
%\item Define a Markov chain according to the network structure, such that from each node a random walker will jump to its neighbors with equal probability, i.e. $P=D^{-1} A$ where $D=\diag(d_i)$ and $d_i = \sum_j A_{ij}$;
%\item Compute its stationary distribution $\pi(i)\sim d_i$. 
%\item Define the source set $V_0=\{1\}$ and the target set $V_1=\{34\}$, compute the committor function
%\[  q(x) = Prob(\mbox{trajectory starting from $x$ hitting $V_1$ before $V_0$}) \]
%by solving the following Dirichlet boundary problem
%\[ (L q)(x) = 0, \ \ \ x\in V_u := V -\{1,34\}, \ \ \ q(1)=0,\ q(34)=1. \]
%Find those edges which contains one node $q(x)\leq 0.5$ and the other node $q(x)\geq 0.5$. Such edge set defines a cut of the graph. 
%\item Compute the effective flux on each edge $(x,y)$ by 
%\[ J^+(x,y) = \max(J(x,y) - J(y,x), 0 ),\]
%where
%\[     J(x,y) = 
%     \begin{cases}
%       \pi(x) (1-q(x)) P_{xy}q(y), & x\neq y;\\
%       0, & \text{otherwise}.
%     \end{cases}
%\]
%\item Compute the transition flux through each node $x\in V$ by 
%\[
%     T(x) = \left\{\begin{array}{ll}
%     \sum_{y \in V} J^+(x,y) ,& x\in V_0 \\
%     \sum_{y\in V} J^+(y,x), & x\in V_1 \\
%     \sum_{y\in V} J^+(x,y) = \sum_{y\in V} J^+(y,x), & x\in V_u 
%     \end{array}\right.
%\]
%\item Visualize your results by plotting a directed graph, with an arrow on each edge indicating the effective flux direction $J^+(x,y)>0$, different color marking the cut set of the graph, and if possible edge/node size in proportion to the size of effective/transition flux.
%\end{enumerate}
%
%A reference can be seen at:
%\begin{itemize}
%\item Weinan E, Jianfeng Lu, and Yuan Yao. {\it The Landscape of Complex Networks: Critical Nodes and A Hierarchical Decomposition}. Methods and Applications of Analysis, special issue in honor of Professor Stanley Osher on his 70th birthday, 20(4):383-404, 2013. \\ arXiv: \url{http://arxiv.org/abs/1204.6376}. \\
%Pdf Link: \url{https://github.com/yao-lab/yao-lab.github.io/blob/master/reference/ELY.MAA13.pdf} 
%\end{itemize}
%
%The following matlab codes implement the transition path analysis and reproduce the results in the paper above:
%
%\url{https://github.com/yao-lab/yao-lab.github.io/blob/master/data/karate_tpt.m}
%
%You may explore more networks, including the two other examples in the paper above:
%
%LAO-binding network (provided by Xuhui Huang): \url{https://github.com/yao-lab/yao-lab.github.io/blob/master/data/lao54.mat}
%
%Les Miserables social network: \url{https://github.com/yao-lab/yao-lab.github.io/blob/master/data/lesmis.mat}, or \url{https://github.com/yao-lab/yao-lab.github.io/blob/master/data/lesmis.txt}

\section{PageRank and Chinese Universities}
